---
title: "Problem Set 3, Econ 211C"
author: "Swati Sharma"
output:
  pdf_document: default
  html_document: default
  html_notebook: default
---

# Question 1

Consider the $ARMA(1,1)$ process
\begin{align}
  Y_t & = 3.2 + 0.86 Y_{t-1} + \varepsilon_t - 1.4 \varepsilon_{t-1}, \,\, \varepsilon_t \sim WN(0,1).
\end{align}
Simulate $n=1010$ observations of this process.

#### Solution:

```{r results='hide'}
n = 1010
cc = 3.2
phi = 0.86
theta = -1.4
eps = rnorm(n,0,1)
y = rep(NA,n)
y[1] = cc + phi*cc/(1-phi)
for(i in 2:n){
  y[i] = cc + phi*y[i-1] + eps[i] + theta*eps[i-1]
}
```

```{r}
plot.ts(y[1:100],  main = paste("First 100 simulated values of the ARMA(1,1)"), ylab = "Y")
```

### a. (10 points)
Use the first 1000 observations from your simulation to estimate an $ARMA(1,1)$ model. Feel free to use the `arima` function. Report your parameter estimates with standard errors and the variance of the residuals. Use your the parameter estimates to compute and report forecasts for $s=1,\ldots,10$. Do not use any pre-packaged functions in `R` -- compute and report the forecasts using basic mathematical operations.

#### Solution:

```{r}
arma11Est = arima(y[1:1000],order=c(1,0,1))
cat("Parameter Estimates: \n")
arma11Est$coef
cat("\n", "Standard Errors: \n")
sqrt(diag(arma11Est$var.coef))
cat("\n", "Variance of Residuals:", var(arma11Est$residuals))

```

```{r results='hide'}
steps = 10
muHat = arma11Est$coef[3]
phiHat = arma11Est$coef[1]
thetaHat = arma11Est$coef[2]
yHat = rep(NA,steps)
yHat[1] =  muHat + phiHat*(y[1000]-muHat) + thetaHat*eps[1000]
for(s in 2:steps){
  yHat[s] = muHat + phiHat*(yHat[s-1]-muHat)
}
```

```{r}
cat("Forecast: \n", yHat)
```


### b. (10 points)
Estimate an $AR(1)$ model using the first 1000 observations from your simulation. Do not use the `arima` function or any other pre-packaged function. Instead, compute the esimates with matrix operations. Use the parameter estimates to compute and report forecasts for $s=1,\ldots,10$. Do not use any pre-packaged functions in `R` -- compute and report the forecasts using basic mathematical operations.

#### Solution:

```{r}
Y = y[1:(n-10)] ## first 1000 simulated values of y
X = cbind(rep(1,n-10),y[2:(n-9)])  ## X matrix of 1000 ones and lagged values of y (t-1)
beta = solve(t(X)%*%X)%*%t(X)%*%Y  ## solving for coefficient
cat("Parameter Estimates: \n", beta)
```
```{r}
steps = 10
phiHatAR = beta[2]
muHatAR = beta[1]/(1-phiHatAR)
yHatAR = rep(NA,steps)
yHatAR[1] =  muHatAR + phiHatAR*(y[1000]-muHatAR) 
yHatAR[2] =  muHatAR + phiHatAR*(yHat[1]-muHatAR)
for(s in 3:steps){
  yHatAR[s] = muHatAR + phiHatAR*(yHatAR[s-1]-muHatAR)
}
cat("Forecast: \n", yHat)
```

### c. (15 points)
Compute the theoretical MSE values for the forecasts associated with $s=1,\ldots,10$, using both the $ARMA(1,1)$ and $AR(1)$ parameter estimates.

#### Solution:

```{r}

## calculating psi coefficients for MA infinite representation of ARMA(1,1)
ARMApsi = ARMAtoMA(arma11Est$coef[1], arma11Est$coef[2], 1)

ar1Est = arima(y[1:1000],order=c(1,0,0))

ARMAmse = matrix(NA,10,1)
ARmse = matrix(NA,10,1)

for (i in 1:10) {
  ## calculating theoretical MSE for ARMA(1,1)
  ARMAmse[i,] = var(arma11Est$residuals)*sum(ARMApsi^(2*i))
  ## calculating theoretical MSE for AR(1)
  ARmse[i,] = var(ar1Est$residuals)*sum(beta[2]^(2*i))
}
cat("Theoretical MSE for forecast using ARMA(1,1) \n", ARMAmse)
cat("\nTheoretical MSE for forecast using AR(1) \n", ARmse)



```

### d. (25 points)
Simulate 100 new datasets of size $n=1010$. For each dataset, estimate the $ARMA(1,1)$ and $AR(1)$ models using the first $1000$ observations and compute the forecasts for $s=1,\ldots,10$. Compute the forecast errors for each set of forecasts (using the true values that you simulated), and compute the sample MSEs of your forecasts.

#### Solution:

```{r error=FALSE, warning=FALSE, message=FALSE}

n = 1010
cc = 3.2
phi = 0.86
theta = -1.4

steps = 10

ARMAmse = matrix(NA, 100, 1)
ARmse = matrix(NA, 100, 1)

for(j in 1:100) {
  ## simulation
  eps = rnorm(n,0,1)
  y = rep(NA,n)
  y[1] = cc + phi*cc/(1-phi)
  for(i in 2:n){
    y[i] = cc + phi*y[i-1] + eps[i] + theta*eps[i-1]
  }
  
  ## ARMA(1,1)
  arma11Est = arima(y[1:1000],order=c(1,0,1))

  muHat = arma11Est$coef[3]
  phiHat = arma11Est$coef[1]
  thetaHat = arma11Est$coef[2]
  
  yHat = rep(NA,steps)
  yHat[1] =  muHat + phiHat*(y[1000]-muHat) + thetaHat*eps[1000]
  for(s in 2:steps){
    yHat[s] = muHat + phiHat*(yHat[s-1]-muHat)
  }
  ARMAmse[j,1] = mean((yHat-y[1000:steps])^2)

  ## AR(1)
  Y = y[1:(n-10)] ## first 1000 simulated values of y
  X = cbind(rep(1,n-10),y[2:(n-9)])  ## X matrix of 1000 ones and lagged values of y (t-1)
  beta = solve(t(X)%*%X)%*%t(X)%*%Y  ## solving for coefficient

  phiHatAR = beta[2]
  muHatAR = beta[1]/(1-phiHatAR)
  yHatAR = rep(NA,steps)
  yHatAR[1] =  muHatAR + phiHatAR*(y[1000]-muHatAR) 
  yHatAR[2] =  muHatAR + phiHatAR*(yHat[1]-muHatAR)
  for(s in 3:steps){
    yHatAR[s] = muHatAR + phiHatAR*(yHatAR[s-1]-muHatAR)
  }
  ARmse[j,1] = mean((yHatAR-y[1000:steps])^2)
}
cat("ARMA(1,1) MSE Values: \n", ARMAmse, "\nAR(1) MSE Values: \n", ARmse)
```

# Question 2
Download daily 1-year and 10-year U.S. Treasury yield data for the period May 21, 2013 - May 20, 2018.

```{r warning=FALSE, message=FALSE}
library(Quandl)
library(vars)
treasury = Quandl("USTREASURY/YIELD",start_date="2013-05-21",end_date="2018-05-20",type="xts")
treasury <-  as.data.frame(treasury)
```

### a. (10 points)
Estimate a bivariate $VAR$ for the log interest rates. Report your parameter estimates and provide some interpretation.

#### Solution:

```{r}
treasury <- treasury[,c(4,9)]
treasury <- log(treasury)
varEst <- VAR(treasury, p=1)
coef(varEst)
```
The estimated VAR suggests that the lagged (by a single time period) values of 1-year and 10-year treasury log interest rates increase the next period's 1-year log rate by 0.9989 and -0.0111 points respectively if they both equal 1. These increases are to a constant of 0.0106. However, the p-value on the lagged 10-year term is quite large.

They also increase the next period's 10-year log rate by -2.51e-05 and 9.93e-01 respectively if they are equal to 1. These increases are to a constant of 5.755e-03. The p-value on the lagged 1-year term is very large. 

### b. (15 points)
Compute and report the impulse response functions of the fitted $VAR$ model.

#### Solution:

```{r}
#varFit <- fitted(varEst)
irf(varEst, n.ahead=15, ortho = FALSE, boot=FALSE)
```

### c. (15 points)
Use the fitted $VAR$ model to compute and report 15-step forecasts for the two series.

#### Solution:

```{r}
predict(varEst, n.ahead=15)
```